\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-12-23 proof-adaptive-optimality.tex
Finished writing major proofs and fixing references. Still needs a once through
corrected various typos

The following lemma is essential to the the proofs
of Theorems~\ref{thm:adpative_lower_bound}, \ref{thm:LAN1}, and
\ref{thm:non_existence}.

@@

  which is bounded from from above by $\eta_\delta(x_{i^*})$ by the induction hypothesis. The case where $N$ is odd and $-x^* \leq y_1 = \ldots = y_N$ is similar. 

@@

  In the case where $c(x)$ is not strictly concave, so that $c'$ does not
  strictly decrease, we approximate $c$ using another concave symmetric
  function with decreasing derivative. We assume w.l.o.g.\ that $c(0) = 0$
  maximizes $c$.  For $\alpha>0$ consider the function $f_\alpha(x) =
  \kappa(\alpha) e^{-|c(x)|^{1+\alpha}}$, where $\kappa(\alpha)$ normalizes
  $f_\alpha$.  Then $c_\alpha(x)$ is concave, symmetric, and
  a.e.\ differentiable with
  \begin{equation*}
    c_\alpha'(x) \triangleq \frac{f'_\alpha(x)}{f_\alpha(x)} = (1+\alpha)|c(x)|^{\alpha} c'(x). 
  \end{equation*}
  Now $c_\alpha'(x)$ is non-increasing since it is the derivative of a concave
  function. Furthermore, since $c(x)$ is non-constant on any interval and
  $c'(x)$ is non-increasing, $c_\alpha'(x)$ is non-constant on any interval
  hence an injection. It follows from the first part of the proof that, for
  any $\alpha>0$,
  \begin{align}
    \label{eq:proof:lem:bound_intervals}
    \frac{(\delta_{n,\alpha})^2}{\Delta_{n,\alpha}(1-\Delta_{n,\alpha})} \leq \max_i \eta_{\alpha}(x_i),
  \end{align}
  where 
  \begin{equation*}
    \delta_{n,\alpha} \triangleq  \sum_{k=1}^{n} (-1)^{k+1} f_{\alpha}(x_k),
  \end{equation*}
  \begin{equation*}
    \Delta_{n,\alpha} \triangleq \sum_{k=1}^n (-1)^{k+1} F_{\alpha}(x_k), 
  \end{equation*}
  and 
  \begin{equation*}
    \eta_{\delta,\alpha}(x) \triangleq \frac{(f_\alpha(x))^{2+\delta}}{\left(F_{\alpha}(x)(1-F(x)) \right)^{1+\delta}}. 
  \end{equation*}
  The proof is completed by noting that 
  \begin{align*}
    \lim_{\alpha \to 0} \frac{(\delta_{n,\alpha})^{2+\delta} }{ \left(\Delta_{n,\alpha}(1-\Delta_{n,\alpha})\right)^{1+\delta}}  = \frac{(\delta_{n})^{2+\delta }}{\left(\Delta_{n}(1-\Delta_{n}) \right)^{1+\delta}},  
  \end{align*}
  and, since the maximum is over a finite set,
  \begin{align*}
    \lim_{\alpha \to 0}  \max_i \eta_{\delta,\alpha}(x_i)  = \max_i\eta_\delta(x_i).
  \end{align*}

>>>>

The following lemma is essential to the proofs
of Theorems~\ref{thm:adpative_lower_bound} and
\ref{thm:non_existence}.

@@

  which is bounded from above by $\eta_\delta(x_{i^*})$ by the induction hypothesis. The case where $N$ is odd and $-x^* \leq y_1 = \ldots = y_N$ is similar. 

@@

  In the case where $c(x)$ is not strictly concave, so that $c'$ does not
  strictly decrease, we approximate $c$ using another concave symmetric
  function with decreasing derivative. We assume w.l.o.g.\ that $c(0) = 0$
  maximizes $c$.  For $\alpha>0$ consider the function $f_\alpha(x) =
  \kappa(\alpha) e^{-|c(x)|^{1+\alpha}}$, where $\kappa(\alpha)$ normalizes
  $f_\alpha$.  Then $c_\alpha(x)$ is concave, symmetric, and
  a.e.\ differentiable with
  \begin{equation*}
    c_\alpha'(x) \triangleq \frac{f'_\alpha(x)}{f_\alpha(x)} = (1+\alpha)|c(x)|^{\alpha} c'(x). 
  \end{equation*}
  Now $c_\alpha'(x)$ is non-increasing since it is the derivative of a concave
  function. Furthermore, since $c(x)$ is non-constant on any interval and
  $c'(x)$ is non-increasing, $c_\alpha'(x)$ is non-constant on any interval
  hence an injection. It follows from the first part of the proof that, for
  any $\alpha>0$,
  \begin{align}
    \label{eq:proof:lem:bound_intervals}
    \frac{(\delta_{n,\alpha})^{2+\delta}}{\left(\Delta_{n,\alpha}(1-\Delta_{n,\alpha})\right)^{1+\delta}} \leq \max_i \eta_{\delta,\alpha}(x_i),
  \end{align}
  where 
  \begin{equation*}
    \delta_{n,\alpha} \triangleq  \sum_{k=1}^{n} (-1)^{k+1} f_{\alpha}(x_k),
  \end{equation*}
  \begin{equation*}
    \Delta_{n,\alpha} \triangleq \sum_{k=1}^n (-1)^{k+1} F_{\alpha}(x_k), 
  \end{equation*}
  and 
  \begin{equation*}
    \eta_{\delta,\alpha}(x) \triangleq \frac{(f_\alpha(x))^{2+\delta}}{\left(F_{\alpha}(x)(1-F(x)) \right)^{1+\delta}}. 
  \end{equation*}
  The proof is completed by noting that 
  \begin{align*}
    \lim_{\alpha \to 0} \frac{(\delta_{n,\alpha})^{2+\delta} }{ \left(\Delta_{n,\alpha}(1-\Delta_{n,\alpha})\right)^{1+\delta}}  = \frac{(\delta_{n})^{2+\delta }}{\left(\Delta_{n}(1-\Delta_{n}) \right)^{1+\delta}},  
  \end{align*}
  and, since the maximum is over a finite set,
  \begin{align*}
    \lim_{\alpha \to 0}  \max_i \eta_{\delta,\alpha}(x_i)  = \max_i\eta_\delta(x_i).
  \end{align*}

### 2019-12-23 proof-lan-distributed.tex
Finished writing major proofs and fixing references. Still needs a once through
corrected various typos

\section{Proof of Theorem~\ref{theorem:non-adaptive-minimax}}

@@

Now, we write the log-likelihood ratio. We have
\begin{align*}
  \sum_{i = 1}^n \log \frac{p_n(B_i)}{p(B_i)}
  & = 2 \sum_{i = 1}^n \log\left(1 + \half W_{n,i}\right) \\
  & = \sum_{i = 1}^n W_{n,i}
  - \frac{1}{4} \sum_{i = 1}^n W_{n,i}^2
  + \half \sum_{i = 1}^n W_{n,i}^2 R(W_{n,i})
\end{align*}
where the remainder $|R(W_{n,i})| \le |W_{n,i}|$ for $|W_{n,i}| \le 1$.
Using the Taylor expansions of $\sqrt{\cdot}$ and
Lemma~\ref{lemma:derivative-bounds}, we have
\begin{align}
  \nonumber
  \half W_{n,i}
  & = \half \score_\theta(B_i)
  \frac{h_n}{\sqrt{n}}
  \pm \left|\frac{h_n^2}{n} \frac{k_i \lip(f)}{p(B_i)}
  + \frac{h_n^2}{n} \score_\theta(B_i)^2
  + \frac{ h_n^4}{n^2} \frac{k_i^2 \lip(f)^2}{p(B_i)^2}\right| \\
  & \in \pm
  C \left|\frac{\sqrt{k_i}}{\sqrt{n} p(B_i)}
  + \frac{k_i}{n p(B_i)}
  + \frac{\sqrt{k_i}}{p(B_i)^2 n}
  + \frac{k_i^2}{p(B_i)^2 n^2}
  \right| \to 0,
  \label{eqn:max-wni-zero}
\end{align}
where $C < \infty$ depends only on $\lip(f)$ and $h_n$ and so is
uniformly bounded;
the limit follows from
and Assumption~\ref{assumption:detection-regions}\eqref{item:finite-intervals}.
Consequently $\max_i W_{n,i} \to 0$, so that
\begin{equation}
  \label{eqn:almost-at-the-end}
  \sum_{i = 1}^n \log \frac{p_n(B_i)}{p(B_i)}
  = \frac{h}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(B_i)
  - \frac{1}{4} \sum_{i = 1}^n \E[\score_\theta(B_i)^2]
  - \frac{1}{4} \sum_{i = 1}^n W_{n,i}^2 + o_P(1).
\end{equation}

It remains to compute $\E[W_{n,i}^2]$. Using the bounds
that $|\score_\theta(B_i)| \le C \sqrt{k_i} / p(B_i)$ from
Lemma~\ref{lemma:derivative-bounds},
the
expansion~\eqref{eqn:max-wni-zero}
yields
\begin{align*}
  \lefteqn{\left|\E\left[W_{n,i}^2 - \frac{h_n^2}{2n} \score_\theta(B_i)^2\right]
    \right|} \\
  &
  \le \frac{C}{n} \left[
    \frac{k_i^{3/2}}{p(A_i)(1 - p(A_i)) \sqrt{n}}
    + \frac{k_i^{3/2}}{p(A_i)^2 (1 - p(A_i))^2 \sqrt{n}}
    + \frac{k_i^2}{p(A_i)(1 - p(A_i))} \frac{1}{n^{3/2}}
    \right] \\
  & \qquad ~ +
  \frac{C}{n}
  \left[\frac{k_i^2}{p(A_i)(1 - p(A_i))} \frac{1}{n}
    + \frac{k_i}{p(A_i)^3(1 - p(A_i))^3} \frac{1}{n}
    + \frac{k_i^4}{p(A_i)^3 (1 - p(A_i))^3} \frac{1}{n^3} \right],
\end{align*}
where $C$ depends only on $h$ and $\lip(f)$.
Thus
\begin{equation*}
  \sum_{i = 1}^n W_{n,i}^2
  = \frac{h_n^2}{n} \sum_{i = 1}^n \score_\theta(B_i)^2
  + o(1),
\end{equation*}
giving the theorem.

@@

>>>>

\section{Proof of Theorem~\ref{theorem:non-adaptive-minimax}}
\label{sec:proof-non-adaptive-minimax}

@@

@@

Now, we write the log-likelihood ratio. We have
\begin{align*}
  \sum_{i = 1}^n \log \frac{p_n(B_i)}{p(B_i)}
  & = 2 \sum_{i = 1}^n \log\left(1 + \half W_{n,i}\right) \\
  & = \sum_{i = 1}^n W_{n,i}
  - \frac{1}{4} \sum_{i = 1}^n W_{n,i}^2
  + \half \sum_{i = 1}^n W_{n,i}^2 R(W_{n,i})
\end{align*}
where the remainder $|R(W_{n,i})| \le |W_{n,i}|$ for $|W_{n,i}| \le 1$.
Using the Taylor expansions of $\sqrt{\cdot}$ and
Lemma~\ref{lemma:derivative-bounds}, we have
\begin{align}
  \nonumber
  \half W_{n,i}
  & = \half \score_\theta(B_i)
  \frac{h_n}{\sqrt{n}}
  \pm \left|\frac{h_n^2}{n} \frac{k_i \lip(f)}{p(B_i)}
  + \frac{h_n^2}{n} \score_\theta(B_i)^2
  + \frac{ h_n^4}{n^2} \frac{k_i^2 \lip(f)^2}{p(B_i)^2}\right| \\
  & \in \pm
  C \left|\frac{\sqrt{k_i}}{\sqrt{n} p(B_i)}
  + \frac{k_i}{n p(B_i)}
  + \frac{\sqrt{k_i}}{p(B_i)^2 n}
  + \frac{k_i^2}{p(B_i)^2 n^2}
  \right| \to 0,
  \label{eqn:max-wni-zero}
\end{align}
where $C < \infty$ depends only on $\lip(f)$ and $h_n$ and so is
uniformly bounded;
the limit follows from Assumption~\ref{assumption:detection-regions}\eqref{item:finite-intervals}.
Consequently $\max_i W_{n,i} \to 0$, so that
\begin{equation}
  \label{eqn:almost-at-the-end}
  \sum_{i = 1}^n \log \frac{p_n(B_i)}{p(B_i)}
  = \frac{h}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(B_i)
  - \frac{1}{4} \sum_{i = 1}^n \E[\score_\theta(B_i)^2]
  - \frac{1}{4} \sum_{i = 1}^n W_{n,i}^2 + o_P(1).
\end{equation}

It remains to compute $\E[W_{n,i}^2]$. Using the bounds
that $|\score_\theta(B_i)| \le C \sqrt{k_i} / p(B_i)$ from
Lemma~\ref{lemma:derivative-bounds},
the
expansion~\eqref{eqn:max-wni-zero}
yields
\begin{align*}
  \lefteqn{\left|\E\left[W_{n,i}^2 - \frac{h_n^2}{2n} \score_\theta(B_i)^2\right]
    \right|} \\
  &
  \le \frac{C}{n} \left[
    \frac{k_i^{3/2}}{p(A_i)(1 - p(A_i)) \sqrt{n}}
    + \frac{k_i^{3/2}}{p(A_i)^2 (1 - p(A_i))^2 \sqrt{n}}
    + \frac{k_i^2}{p(A_i)(1 - p(A_i))} \frac{1}{n^{3/2}}
    \right] \\
  & \qquad ~ +
  \frac{C}{n}
  \left[\frac{k_i^2}{p(A_i)(1 - p(A_i))} \frac{1}{n}
    + \frac{k_i}{p(A_i)^3(1 - p(A_i))^3} \frac{1}{n}
    + \frac{k_i^4}{p(A_i)^3 (1 - p(A_i))^3} \frac{1}{n^3} \right],
\end{align*}
where $C$ depends only on $h$ and $\lip(f)$.
Thus
\begin{equation*}
  \sum_{i = 1}^n W_{n,i}^2
  = \frac{h_n^2}{n} \sum_{i = 1}^n \score_\theta(B_i)^2
  + o(1),
\end{equation*}
giving Lemma~\ref{lemma:lan-bits}.

### 2019-12-23 proof-sgd.tex
corrected various typos

This proof requires somewhat more technicality than the first part of the
theorem, including a brief detour into local asymptotic normality theory,
regular estimators, and quadratic-mean
differentiability~\cite[cf.]{VanDerVaart98}.
We assume without loss of generality that the median of the density
$f$ is 0, so that if $P_\theta$ has density $f(\cdot - \theta)$, the
median of $P_\theta$ is $\theta$.
We begin by recalling the statistical concepts we require.
\begin{definition}
  \label{definition:regular-estimator}
  A sequence of estimators $T_n$ for a parameter $\theta$ in the parametric
  family $\{P_\theta\}_{\theta \in \Theta}$ is \emph{regular at $\theta$} if
  there exists a distribution $Q$ such that for all bounded sequence
  $h_n$,
  \begin{equation*}
    \sqrt{n}(T_n - (\theta + h_n / \sqrt{n}))
    \mathop{\cd}_{P_{\theta + h_n/\sqrt{n}}}
    Q.
  \end{equation*}
\end{definition}
\begin{definition}
  \label{definition:qmd}
  Let $\{P_\theta\}_{\theta \in \Theta}$ have densities $p_\theta$
  with respect to a base measure $\mu$. The family
  is \emph{quadratic mean differentiable} (QMD) at
  $\theta$ with score $\score_\theta$ if
  \begin{equation}
    \label{eqn:qmd}
    \int \left(\sqrt{p_{\theta + h}} - \sqrt{p_\theta}
    - \half h^\top \score_\theta\sqrt{p_\theta} \right)^2 d\mu
    = o(\norm{h}^2)
  \end{equation}
  as $h \to 0$.
\end{definition}
\begin{definition}
  \label{definition:lan}
  A family of distributions $\{P_\theta\}_{\theta \in \Theta}$ is
  \emph{locally asymptotically normal with information matrix $I_\theta$} (LAN)
  at
  $\theta$ if there exists a sequence of random vectors $Z_n$ such that for
  all $h_n \to h$,
  \begin{equation*}
    \sum_{i = 1}^n \log \frac{dP_{\theta + h_n/\sqrt{n}}}{dP_\theta}
    (X_1, \ldots, X_n)
    = h^\top Z_n - \half h^\top I_\theta h + o_P(1)
  \end{equation*}
  where $Z_n \cd \normal(0, I_\theta)$ under $P_\theta$, where
  $X_i \simiid P_\theta$.
\end{definition}

@@

The regularity result~\eqref{eq:normal_expansion_lem} gives
\begin{equation*}
  \sqrt{n} \bar{\Delta}_n = -\frac{1}{\sqrt{n}}
  \sum_{i = 1}^n  \frac{\varphi(Z_i)}{\psi'(0)} + o_{P,n}(1).
\end{equation*}
The conditions in
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam} imply
that the Fisher information $I_h =
\E_h[\score_h(Z)^2]$ exists and is continuous for
$\score_h(z) = \frac{p'(z - h)}{p(z - h)}$,
and the asymptotic expansion Definitions~\ref{definition:qmd}
and~\ref{definition:lan} allow combines
with the preceding display to give the joint convergence
\begin{equation*}
  \left(\sqrt{n} \bar{\Delta}_n, \log \frac{P^n_{h_n/\sqrt{n}}}{P^n_0}(Z_1^n)
  \right)
  \cd \normal \left( \left(0,-\frac{h^2}{2} I_0 \right),
  \begin{pmatrix}
    \frac{\chi(0)}{\psi'(0)^2} & \frac{-h}{ \psi'(0)}
    \E_p[\varphi(Z) \score_0(Z)] \\
    \frac{-h}{ \psi'(0)} \E_p[\varphi(Z) \score_0(Z)]
    & h^2 I_0
  \end{pmatrix}  \right).
\end{equation*}
Le Cam's third lemma \cite[Exm. 6.7]{VanDerVaart98} then implies the
convergence
\begin{equation*}
  \sqrt{n}\bar{\Delta}_n
  \mathop{\cd}_{P_{h_n/\sqrt{n}}^n}
  \Ncal\left(\frac{-h}{ \psi'(0)} \E_p[\varphi(Z) \score(Z)],
  \frac{\chi(0)}{\psi'(0)^2} \right)
\end{equation*}
under the alternatives $P_{h_n/\sqrt{n}}$, which gives
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam}.

>>>>

This proof requires somewhat more technicality than the first part of the
theorem, including a brief detour into local asymptotic normality theory,
regular estimators, and quadratic-mean
differentiability~\cite[cf.]{VanDerVaart98}.
We assume without loss of generality that the median of the density
$f$ is 0, so that if $P_\theta$ has density $f(\cdot - \theta)$, the
median of $P_\theta$ is $\theta$.
We begin by recalling the statistical concepts we require.
\begin{definition}
  \label{definition:regular-estimator}
  A sequence of estimators $T_n$ for a parameter $\theta$ in the parametric
  family $\{P_\theta\}_{\theta \in \Theta}$ is \emph{regular at $\theta$} if
  there exists a distribution $Q$ such that for any bounded sequence
  $h_n$,
  \begin{equation*}
    \sqrt{n}(T_n - (\theta + h_n / \sqrt{n}))
    \mathop{\cd}_{P_{\theta + h_n/\sqrt{n}}}
    Q.
  \end{equation*}
\end{definition}
\begin{definition}
  \label{definition:qmd}
  Let $\{P_\theta\}_{\theta \in \Theta}$ have densities $p_\theta$
  with respect to a base measure $\mu$. The family
  is \emph{quadratic mean differentiable} (QMD) at
  $\theta$ with score $\score_\theta$ if
  \begin{equation}
    \label{eqn:qmd}
    \int \left(\sqrt{p_{\theta + h}} - \sqrt{p_\theta}
    - \half h^\top \score_\theta\sqrt{p_\theta} \right)^2 d\mu
    = o(\norm{h}^2)
  \end{equation}
  as $h \to 0$.
\end{definition}
\begin{definition}
  \label{definition:lan}
  A family of distributions $\{P_\theta\}_{\theta \in \Theta}$ is
  \emph{locally asymptotically normal with information matrix $I_\theta$} (LAN)
  at
  $\theta$ if there exists a sequence of random vectors $Z_n$ such that for
  all $h_n \to h$,
  \begin{equation*}
    \sum_{i = 1}^n \log \frac{dP_{\theta + h_n/\sqrt{n}}}{dP_\theta}
    (X_1, \ldots, X_n)
    = h^\top Z_n - \half h^\top I_\theta h + o_P(1)
  \end{equation*}
  where $Z_n \cd \normal(0, I_\theta)$ under $P_\theta$, where
  $X_i \simiid P_\theta$.
\end{definition}

@@

The regularity result~\eqref{eq:normal_expansion_lem} gives
\begin{equation*}
  \sqrt{n} \bar{\Delta}_n = -\frac{1}{\sqrt{n}}
  \sum_{i = 1}^n  \frac{\varphi(Z_i)}{\psi'(0)} + o_{P,n}(1).
\end{equation*}
The conditions in
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam} imply
that the Fisher information $I_h =
\E_h[\score_h(Z)^2]$ exists and is continuous for
$\score_h(z) = \frac{p'(z - h)}{p(z - h)}$,
and the asymptotic expansion Definitions~\ref{definition:qmd}
and~\ref{definition:lan} combined
with the preceding display, give the joint convergence
\begin{equation*}
  \left(\sqrt{n} \bar{\Delta}_n, \log \frac{P^n_{h_n/\sqrt{n}}}{P^n_0}(Z_1^n)
  \right)
  \cd \normal \left( \left(0,-\frac{h^2}{2} I_0 \right),
  \begin{pmatrix}
    \frac{\chi(0)}{\psi'(0)^2} & \frac{-h}{ \psi'(0)}
    \E_p[\varphi(Z) \score_0(Z)] \\
    \frac{-h}{ \psi'(0)} \E_p[\varphi(Z) \score_0(Z)]
    & h^2 I_0
  \end{pmatrix}  \right).
\end{equation*}
Le Cam's third lemma \cite[Exm. 6.7]{VanDerVaart98} then implies the
convergence
\begin{equation*}
  \sqrt{n}\bar{\Delta}_n
  \mathop{\cd}_{P_{h_n/\sqrt{n}}^n}
  \Ncal\left(\frac{-h}{ \psi'(0)} \E_p[\varphi(Z) \score(Z)],
  \frac{\chi(0)}{\psi'(0)^2} \right)
\end{equation*}
under the alternatives $P_{h_n/\sqrt{n}}$, which gives
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam}.
